---
title: "Project"
author: "Emily"
date: "9/30/2020"
output: html_document
---
This data is about flu cases/hospitalizations/deaths per state and percent of people who recieved the flu vaccination 18+ years in 2009-2010 Swine Flu Epidemic.

```{r echo=FALSE, results='hide'}
library(tidyverse)
library(ggplot2)
getwd()
setwd("/Users/emilyreed/Desktop")
flu_cases <- read.csv("Flu_cases.csv") 
vaccinations <- read.csv("Vaccinations.csv")
```

Flu cases has 56 observations (1 per state, plus a few territories). "cases", "hospitalizations", and "deaths" refers to the number of cases, hospital visits, and death due to influenza reported in that state for the 2009-2010 flu season. This data is NOT separated by age.

```{r}
#flu cases table
#turning cases and hospitalizations into numeric values for later calculations
flu_cases1 <- flu_cases %>% mutate_at("Hospitalizations", as.numeric) 
glimpse(flu_cases1)
head(flu_cases1)

```

Vaccinations is a dataset with 51 observations (one per state, plus District of Colubmbia). "Perc_vaccinated" refers to the percentage of each age group vaccinated in each state in the 2009-2010 flu season. 

```{r}
#flu vaccinations table
glimpse(vaccinations)
head(vaccinations)
```
Vaccinations dataset needs to be 'tidy'ed!


```{r}
vaccinations2 <- vaccinations %>% 
  pivot_longer(2:4, names_to="age_group",values_to="percentage_vaccinated") %>% 
  separate("age_group", into=c("percent", "age_group"), sep="\\.", convert=T ) %>%
  select(-percent) %>% rename("State"="Names")

vaccinations2
```


```{r}
#Lets join the data!
cases_vaccinations <- left_join(flu_cases1, vaccinations2, by="State", convert=T)
glimpse(cases_vaccinations)
head(cases_vaccinations)
```
-----
Using summary statistics:


#Finding the average % vaccinated for each state across all age groups

```{r}
#Find mean % of vaccinations across all age groups for each state
cases_vaccinations <- cases_vaccinations %>% group_by(State) %>% mutate(average_vaccinations= mean(percentage_vaccinated))
cases_vaccinations

#find states with average vaccinations % that are above the National average vaccination %
national_average_vaccinated <- cases_vaccinations %>% filter(!is.na(average_vaccinations)) %>%  summarize(national_avg=mean(average_vaccinations)) %>% summarize(mean(national_avg)) 
national_average_vaccinated


cases_vaccinations%>% filter(average_vaccinations>national_average_vaccinated) %>% select(State, average_vaccinations) %>% summarize(avg_vaccination_percent = mean(average_vaccinations)) %>% arrange(desc(avg_vaccination_percent))


```
##Lets see which states had the most cases of flu. Which states had the fewest percent average vaccinated citizens?

```{r}
#Which states had the most cases of flu
cases_vaccinations %>% 
  group_by(State) %>% 
  summarize(case=mean(cases)) %>% 
  arrange(desc(case))


#Which states had the lowest average % vaccinated across all age groups
cases_vaccinations %>% 
  group_by(State) %>%
  summarize(avg_vaccination_percent = mean(average_vaccinations)) %>% 
  arrange(avg_vaccination_percent)


#Do the states that had the most flu match with the states that had the lowest average % vaccinated? (Nope)
cases_vaccinations %>% 
  group_by(State) %>% 
  summarize(case_in_state=mean(cases), avg_vaccination_state=(mean(average_vaccinations))) %>% 
  arrange(desc(case_in_state), avg_vaccination_state) %>% 
  select(State, case_in_state, avg_vaccination_state)

```


##Make new variables using mutate for mean and sd for each % vaccinated in each age group average cases, hospitalizations and deaths

Finding the national average of cases
```{r}
national_average_cases <- cases_vaccinations %>% summarize(average_cases_US= mean(cases)) %>% summarize(average_cases_US=mean(average_cases_US))
national_average_cases
```


Finding the national sd of cases between states
```{r}
national_sd_cases <- cases_vaccinations %>% summarize(sd_cases_US= mean(cases)) %>% summarize(sd_cases=sd(sd_cases_US))
national_sd_cases
```


Finding the national average of hospitalizations
```{r}
national_average_hospitalizations <- cases_vaccinations %>% 
  filter(!is.na(Hospitalizations)) %>% 
  summarize(average_hosp_US= mean(Hospitalizations)) %>%
  summarize(average_hosp_US=mean(average_hosp_US))
national_average_hospitalizations
```


Finding the national sd of hospitalizations between states
```{r}
national_sd_hospitalizations<- cases_vaccinations %>% 
  filter(!is.na(Hospitalizations)) %>% 
  summarize(sd_hosp_US= mean(cases)) %>% 
  summarize(sd_hosp_US=sd(sd_hosp_US))
national_sd_hospitalizations
```


Finding the national average of deaths (this data is obviously not very accurate, oop)
```{r}
national_average_deaths <- cases_vaccinations %>% 
  filter(!is.na(Deaths)) %>% 
  summarize(average_deaths_US= mean(Deaths)) %>%
  summarize(average_deaths_US=mean(average_deaths_US))
national_average_deaths
```


Finding the national sd of deaths between states
```{r}
national_sd_deaths<- cases_vaccinations %>% 
  filter(!is.na(Deaths)) %>% 
  summarize(sd_deaths_US= mean(cases)) %>% 
  summarize(sd_deaths_US=sd(sd_deaths_US))
national_sd_deaths
```



#Find quantile, min, max for all numeric variables (make into table format using ktables)
```{r}


```


#make a correlation heat map of numeric variables
```{r}

```



#Average vaccination % (for all age groups) and cases per state graph (scatterplot)

```{r}

cases_vaccinations %>% 
  group_by(State) %>% 
  filter(!is.na(average_vaccinations)) %>% 
  mutate(avg_vaccination_percent = mean(average_vaccinations)) %>% 
  select(State, avg_vaccination_percent, cases, age_group) %>% 
  filter(cases > national_average_cases) %>% 
  ggplot()+
  geom_point(aes(avg_vaccination_percent, cases, col=State))+
  ggtitle("Average % Vaccinations and Cases for State Cases Above National Average")+
  xlab("Average Vaccination Percent")+
  ylab("Number of Cases")

```




States vaccination percentage for each age group, facet by state, summary stat????

#want to color facet title whether they were over national average number of cases or by quantile, or under and fix order of x variables

```{r , fig.height=10, fig.width=13}

cases_vaccinations %>% 
  group_by(State) %>% 
  filter(!is.na(percentage_vaccinated)) %>% 
  ggplot(aes(x=age_group ,y=percentage_vaccinated, fill=age_group))+
  geom_bar(stat="identity") +
  facet_wrap(~State) +
  geom_hline(yintercept = 51.62267, linetype="longdash")+
  theme(axis.title.x=element_blank(), axis.text.x=element_blank())+
  ylab("Percent Vaccinated") 
  
#filter(case_vaccinations$case>national_average_cases) %>% 
  #theme(element_rect(fill="red"))+

```

